---
title: "R_techniques"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Speedup R code for data manipulation and visualization

## Vectorization

```{r}
vec <- seq(1, 200, 1)
# add 10 to each element of the vector
for(i in 1:length(vec)){
  vec[i] <- vec[i] + 10
}

vec <- seq(1, 200, 1)
# add 10 to each element of the vector (use vectorization)
vec <- vec + 10
```

## Built-in function > for loop > recursion

```{r}
vec <- seq(1, 1e8, 1)
# Built-in function
system.time(
  sum(vec)
)

# For loop
custom_sum <- function(para){
  sum <- 0
  for (i in vec){
    sum = sum + i
  }
  return(sum)
}
system.time(custom_sum(vec))
```

```{r}
N = 30

# For loop (Fibonacci sequence)
fibo_for <- function(N){
  fibo_seq <- rep(0, N)
  fibo_seq[1] <- 0
  fibo_seq[2] <- 1
  for (i in 1:N){
    if (i >= 3){
      fibo_seq[i] = fibo_seq[i - 1] + fibo_seq[i - 2]
    }
  }
  print(fibo_seq[N])
}
system.time(fibo_for(N))

# Recursion (Fibonacci sequence)
fibo_recur <- function(k){
  if (k == 1) {
    return(0)
  } else if (k == 2) {
    return(1)
  } else {
    return(fibo_recur(k-1) + fibo_recur(k-2))
  }
}
system.time(
  print(fibo_recur(N))
)
```

## apply(), lapply(), sapply(), tapply()

```{r}
# a 100 column data.frame
x <- data.frame(matrix(runif(100*1e4), ncol = 100))

# loop 100 times, adding another row to x
system.time(
  for(i in seq_along(1:100)){
    x <- rbind(x, data.frame(matrix(runif(1*1e4), ncol = 100)))
    }
)
```

```{r}
# a 100 column data.frame
x <- data.frame(matrix(runif(100*1e4), ncol = 100))

## using lapply()
system.time(
  lapply(1:100, function(y) rbind(x, data.frame(matrix(runif(1*1e4), ncol = 100))))

)
```

## Parallel Processing

Parallel computing involves the simultaneous execution of multiple tasks or processes to speed up computations, where we can take advantage of multi-core processors. In R, we can leverage many packages to perform parallel computing. 


# Example on Real Data

```{r cars}
# load libraries
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

```{r}
accident = read.csv("accident_subset.csv")
```

## for loop
```{r}
unique_states <- unique(accident$State)
system.time(
for(state in unique_states){
  # Subset data for the current state
  state_data <- accident[accident$State == state,]
  
  p <- ggplot(state_data, aes(x=Severity)) +
    geom_histogram(binwidth=1, fill="blue", color="black", alpha=0.7) + 
    ggtitle(paste("Distribution of Severity in", state)) +
    xlab("Severity") +
    ylab("Count") +
    theme_minimal()
  
  print(p)
})
```


## apply() + Parallel Processing
```{r}
library(parallel)
plot_histogram <- function(state) {
  # Subset data for the current state
  state_data <- accident[accident$State == state,]
  
  # Plot histogram
  p <- ggplot(state_data, aes(x=Severity)) +
    geom_histogram(binwidth=1, fill="blue", color="black", alpha=0.7) + 
    ggtitle(paste("Distribution of Severity in", state)) +
    xlab("Severity") +
    ylab("Count") +
    theme_minimal()
  
  print(p)

}

# Use mclapply with the number of cores
# detectCores() can get the number of available cores in the system
system.time(
  mclapply(unique_states, plot_histogram, mc.cores = detectCores())
)
```


